---
title: "Effects of subscription programme"
output: pdf_document
---

```{r setup, include=FALSE}
#output_format: "word_document"  pdf_document
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=3.2, fig.height=3.5) 

options(scipen = 999) # disable scientific notation (this causes long numbers to appear in full rather than being truncated)
# Seed for random number generation
set.seed(42)

#library("readxl")
library("dplyr")
library("ggplot2")
library("effects")
library("DT")
library("readxl")


source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
source("http://www-users.york.ac.uk/~er13/45I%20-%202016/functions/summarySE.R")
```

*Please note that the analysis below was ran on a simulated dataset and should be only considered in the context of the portfolio.*

# Question: What is  the impact of joining the subscription programme on a business that cares about customer ordering frequency and value?

## Executive summary

* The subscription plan significantly increases the frequency and value of orders in the whole sample when compared to the pre-subscription stage. 
* The highest increases appear during the active subscription stage, but there are carry-over increases in the measures after cancellation, suggesting behavioural changes in the customers. 
* The analysis revealed further differences between customers who retain their active subscription status and those who cancel around the end of the trial phase of the programme. Despite making up only 24 % of the whole sample the Retention group represents very valuable customers as they place significantly more valuable orders more frequently than the Churn group. 
* The increases in the frequency and value of orders associated with the subscription programme translate into a clear growth of revenue. Within the Retention group we can see a linear increase in the frequency and value of orders with time, specifically when entering their fourth active subscription month.

## Data analysis

```{r include=FALSE}

#explore means, sd, frequencies in general

subscr_data<-read_excel("my_data.xlsx")
head(subscr_data)
no_eaters<-length(unique(subscr_data$SUBSCRIBER_ID))
no_months<-length(unique(subscr_data$MONTH_START))

subscr_data$SUBSCRIPTION_STATUS_IN_MONTH <- factor( subscr_data$SUBSCRIPTION_STATUS_IN_MONTH, levels = c("pre subscription", "active","cancelled"))

#check for NA-s in factor 
sum(is.na(subscr_data$SUBSCRIPTION_STATUS_IN_MONTH))>0
#careful is plyr and dplyr are both loaded summarise will give an error
sum_month_data<-subscr_data %>%
  group_by(SUBSCRIPTION_STATUS_IN_MONTH) %>%
  summarise(n_obs = n(),
            m_ORDERS = mean(ORDERS),
            sd_ORDERS = sd(ORDERS_VALUE),
            m_VALUE = mean(ORDERS_VALUE),
            sd_VALUE = sd(ORDERS_VALUE))

sum_current_data<-subscr_data %>%
  group_by(CURRENT_STATUS) %>%
  summarise(n_obs = n(),
            m_ORDERS = mean(ORDERS),
            sd_ORDERS = sd(ORDERS_VALUE),
            m_VALUE = mean(ORDERS_VALUE),
            sd_VALUE = sd(ORDERS_VALUE))
```

One can re-frame the above question on a data analysis level in the following way: Does the monthly subscription stage affect ordering frequency and value?

All-analyses were conducted using *R* (Version 3.5.1). For the analysis script and exact test statistics please refer to the corresponding R MarkDown document. 

When exploring the data set, we see that it consists of different numbers of observation from N =`r no_eaters` customers over `r no_months` months of uninterrupted time-period. The descriptive statistics of the data set (see table below) reveal that on average, customers seem to increase both during an active subscription phase and after cancelling the programme. However, one needs to run test statistics to see if the different subscription states are statistically different from one another. 

```{r echo=FALSE}
knitr::kable(sum_month_data, caption = "Summary of average ordering frequency and value based on the monthly subscription status", col.names = c("Subscription status", "Observations", "Order frequency", "SD order frequency", "Order value", "SD value order"), digits=2)
```

It is good practice to filter out outliers (here defined as values +/- 3 standard deviations away from the mean) before running statistical analyses, but here the data set remains the same after filtering.

```{r include=FALSE}

data_outlier_max<-mean(subscr_data$ORDERS_VALUE)+3*sd(subscr_data$ORDERS_VALUE)
data_outlier_min<-mean(subscr_data$ORDERS_VALUE)-3*sd(subscr_data$ORDERS_VALUE)

data_outlier_max_orders<-mean(subscr_data$ORDERS)+3*sd(subscr_data$ORDERS)
data_outlier_min_orders<-mean(subscr_data$ORDERS)-3*sd(subscr_data$ORDERS)

rm_subscr_data <- subscr_data %>%
  filter(ORDERS_VALUE < data_outlier_max,ORDERS_VALUE > data_outlier_min)%>%
  filter(ORDERS < data_outlier_max_orders,ORDERS > data_outlier_min_orders)

no_rm_eaters<-length(unique(rm_subscr_data$SUBSCRIBER_ID))
```

Nonetheless, when plotting the distribution and individual data points with their central metrics (see figures below), it is clear that the distribution skewed (frequently the case with real-world data) and the frequency of 0 values are overrepresented in the sample. To address this, further filtering and transformations (or fitting models with non-normal distributions) could be done, but for this exercise I decided to keep the untransformed values as they are. 

```{r echo=FALSE}

#plot distribution and central metrics
ggplot(rm_subscr_data, aes(x = factor(SUBSCRIPTION_STATUS_IN_MONTH), y = ORDERS_VALUE, fill = SUBSCRIPTION_STATUS_IN_MONTH)) +
  geom_flat_violin(aes(fill = SUBSCRIPTION_STATUS_IN_MONTH),position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA)+
  geom_point(aes(x = as.numeric(SUBSCRIPTION_STATUS_IN_MONTH)-.15, y = ORDERS_VALUE, colour = SUBSCRIPTION_STATUS_IN_MONTH),position = position_jitter(width = .05), size = .25, shape = 20,alpha = 0.1)+
  geom_boxplot(aes(x = as.numeric(SUBSCRIPTION_STATUS_IN_MONTH), y = ORDERS_VALUE, fill = SUBSCRIPTION_STATUS_IN_MONTH),outlier.shape = NA, alpha = .5, width = .1, colour = "black")+ theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))+xlab("Subscription stage")+ylab("")+coord_flip()+ggtitle("Value of orders")

ggplot(rm_subscr_data, aes(x = factor(SUBSCRIPTION_STATUS_IN_MONTH), y = ORDERS, fill = SUBSCRIPTION_STATUS_IN_MONTH)) +
  geom_flat_violin(aes(fill = SUBSCRIPTION_STATUS_IN_MONTH),position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA)+
  geom_point(aes(x = as.numeric(SUBSCRIPTION_STATUS_IN_MONTH)-.15, y = ORDERS, colour = SUBSCRIPTION_STATUS_IN_MONTH),position = position_jitter(width = .05), size = .25, shape = 20,alpha = 0.1)+
  geom_boxplot(aes(x = as.numeric(SUBSCRIPTION_STATUS_IN_MONTH), y = ORDERS, fill = SUBSCRIPTION_STATUS_IN_MONTH),outlier.shape = NA, alpha = .5, width = .1, colour = "black")+ theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))+xlab("")+ylab("")+scale_x_discrete(
                     labels = c("","",""))+coord_flip()+ggtitle("Frequency of orders")
```

Given that monthly subscription state has 3 levels, linear regression was applied to see its effects on order frequency and value.

```{r echo=FALSE}

# statistics of the whole sample on the effects of SUBSCRIPTION STATUS
orders_mod <- lm(ORDERS ~ SUBSCRIPTION_STATUS_IN_MONTH, data = rm_subscr_data)
value_mod <- lm(ORDERS_VALUE ~ SUBSCRIPTION_STATUS_IN_MONTH, data = rm_subscr_data)
#gather stats
orders_mod_res<-summary(orders_mod)
value_mod_res<-summary(value_mod)
```

The analysis revealed a significant increase from pre-subscription stage both in the active and cancelled states for the frequency and value of the orders (see figure below). This result suggests that a subscription has facilitating effect on ordering behaviour, even after cancelling the programme. Both fitted models explained `r round(orders_mod_res$adj.r.squared,2)*100`% of the variance for both outcome measures.

```{r echo=FALSE}
#plot estimates
layout(matrix(c(1,2), 1, 2, byrow = TRUE))
plot(allEffects(orders_mod), ylab = "Frequency of order", xlab = "Status", main = NULL)
plot(allEffects(value_mod),ylab = "Value of order", xlab = "Status",main = NULL)

```

While the results seem to suggest the efficiency of the subscription programme the question arises whether this is true for all types of customers. It is possible that customers who were only interested in the two-week-long, free subscription period are different from the customers who retained their subscription.

```{r include=FALSE}

# analysis on two types of eaters
#create dataset 
eaters_data <- subscr_data %>%
  group_by(SUBSCRIBER_ID,SUBSCRIPTION_STATUS_IN_MONTH) %>%
  summarise( n = n(),
             m_ORDERS = mean(ORDERS),
            sd_ORDERS = sd(ORDERS_VALUE),
            m_VALUE = mean(ORDERS_VALUE),
            sd_VALUE = sd(ORDERS_VALUE))
#active for more than 1 month - to avoid people who accidentally stayed subscribed
ret_eaters<-eaters_data %>%
  filter(n>1, SUBSCRIPTION_STATUS_IN_MONTH == "active")
ret_eaters_ID<-ret_eaters$SUBSCRIBER_ID

ret_eaters_month <- subscr_data[subscr_data$SUBSCRIBER_ID %in% ret_eaters_ID,]
free_eaters_month <- subscr_data[!subscr_data$SUBSCRIBER_ID %in% ret_eaters_ID,]
#prepare reference SD to exclude outliers in both groups and in both measures 
outlier_max<-mean(ret_eaters_month$ORDERS_VALUE)+3*sd(ret_eaters_month$ORDERS_VALUE)
outlier_min<-mean(ret_eaters_month$ORDERS_VALUE)-3*sd(ret_eaters_month$ORDERS_VALUE)

outlier_max_orders<-mean(ret_eaters_month$ORDERS)+3*sd(ret_eaters_month$ORDERS)
outlier_min_orders<-mean(ret_eaters_month$ORDERS)-3*sd(ret_eaters_month$ORDERS)

f_outlier_max<-mean(free_eaters_month$ORDERS_VALUE)+3*sd(free_eaters_month$ORDERS_VALUE)
f_outlier_min<-mean(free_eaters_month$ORDERS_VALUE)-3*sd(free_eaters_month$ORDERS_VALUE)

f_outlier_max_orders<-mean(free_eaters_month$ORDERS)+3*sd(free_eaters_month$ORDERS)
f_outlier_min_orders<-mean(free_eaters_month$ORDERS)-3*sd(free_eaters_month$ORDERS)

```

At this stage customers were split into two groups such as the Retention group and Churn group based on the number of active monthly subscriptions. To avoid the inclusion of accidental customers who might have forgotten to cancel, only customers with more than 1 active month were included in the Retention group. The rest of the customers made up the Churn group. 

```{r include=FALSE}
#remove outliers from RETENTION group
rm_ret_eaters_month <- ret_eaters_month %>%
  filter(ORDERS_VALUE < outlier_max,ORDERS_VALUE > outlier_min)%>%
  filter(ORDERS < outlier_max_orders,ORDERS > outlier_min_orders)

#looks like whoever had 1 active status will also have more than 1
sum_ret<- rm_ret_eaters_month %>%
  group_by(SUBSCRIPTION_STATUS_IN_MONTH) %>%
  summarise( n = n(),
             m_ORDERS = mean(ORDERS),
            sd_ORDERS = sd(ORDERS_VALUE),
            m_VALUE = mean(ORDERS_VALUE),
            sd_VALUE = sd(ORDERS_VALUE))

#remove outliers from CHURN group
rm_free_eaters_month <- free_eaters_month %>%
  filter(ORDERS_VALUE < f_outlier_max,ORDERS_VALUE > f_outlier_min)%>%
  filter(ORDERS < f_outlier_max_orders,ORDERS > f_outlier_min_orders)

#people here cancelled in the same month - probably only there for the 2 week period
sum_free<- rm_free_eaters_month %>%
  group_by(SUBSCRIPTION_STATUS_IN_MONTH) %>%
  summarise( n = n(),
             m_ORDERS = mean(ORDERS),
            sd_ORDERS = sd(ORDERS_VALUE),
            m_VALUE = mean(ORDERS_VALUE),
            sd_VALUE = sd(ORDERS_VALUE))

```

The summary tables (see below) suggest that customers who had at least 1 active monthly subscription also kept their subscription for 1 other month at least. Specifically, this can be seen in Churn group where there was no-one in the Active stage throughout the 10 month period. We can see that both the frequency and the value of the average order per month increased during the active months for the Retention group. Interestingly, even after cancelling the subscription (i.e. after an active period or around the 2 week-long trial) the frequency and value seem to remain higher than in the pre-subscription period - suggesting behavioural change in the customer. Another, interesting observation is that the Retention and Churn group differ in terms of their baseline (i.e. pre-subscription). This data could be a useful starting point when preparing a strategy for targeting certain customers (but it is not under the scope of the present analysis).

```{r echo=TRUE, echo=FALSE}
knitr::kable(sum_ret, caption = "Summary of average ordering behaviour for the Retention group", col.names = c("Subscription status", "Observations", "Order frequency", "SD order frequency", "Order value", "SD value order"), digits=2)
```


```{r echo=TRUE, echo=FALSE}

knitr::kable(sum_free, caption = "Summary of average ordering behaviour for the Churn group", col.names = c("Subscription status", "Observations", "Order frequency", "SD order frequency", "Order value", "SD value order"),digits=2)
```

To test the effects of the monthly subscription state within the two groups separately, linear regressions were run on the data. The analysis revealed the significant increasing effect of monthly subscription state on both order value and frequency in both groups (to a higher extent in the Retention group).

```{r include=FALSE}

#regression on the retention group w/o outliers
ret.mod.order<-lm(ORDERS ~ SUBSCRIPTION_STATUS_IN_MONTH, data = rm_ret_eaters_month)
ret.mod.value<-lm(ORDERS_VALUE ~ SUBSCRIPTION_STATUS_IN_MONTH, data = rm_ret_eaters_month)
#gather statistics
summary(ret.mod.order)
summary(ret.mod.value)

#regression on the retention group w/o outliers
free.mod.order<-lm(ORDERS ~ SUBSCRIPTION_STATUS_IN_MONTH, data = rm_free_eaters_month)
free.mod.value<-lm(ORDERS_VALUE ~ SUBSCRIPTION_STATUS_IN_MONTH, data = rm_free_eaters_month)
#gather statistics
summary(free.mod.order)
summary(free.mod.value)
```

```{r echo=FALSE}
#plot effects (using estimates and conf.intervals)
plot(allEffects(ret.mod.order),ylab = "Frequency of orders", xlab = "Status",main = "Retention group")
plot(allEffects(ret.mod.value), ylab = "Value of orders", xlab = "Status",main = "Retention group")
```

```{r echo=FALSE}
#plot effects (using estimates and conf.intervals)
plot(allEffects(free.mod.order), ylab = "Frequency of orders", xlab = "Status",main = "Churn group")
plot(allEffects(free.mod.value),ylab = "Value of orders", xlab = "Status",main = "Churn group")

```

To see if the two groups of customers were statistically different from one to another two linear regressions were run after combining the two groups into one, using a newly created categorical factor (i.e. group membership: Retention, Churn). Based on the results there is indeed a significant difference between these two kinds of customer at both outcome measures (see figure below).

```{r include=FALSE}

###do the two eater types differ from one-another statistically?
#prepare new factor EATER_TYPE
rm_ret_eaters_month$EATER_TYPE<-"retention"
rm_free_eaters_month$EATER_TYPE<-"churn"
#combine datasets
rm_total <- rbind(rm_ret_eaters_month, rm_free_eaters_month)
rm_total$EATER_TYPE<- factor(rm_total$EATER_TYPE)

#test their difference on the combined dataset
rm_total.mod.order<-lm(ORDERS ~ EATER_TYPE, data = rm_total)
rm_total.mod.value<-lm(ORDERS_VALUE ~ EATER_TYPE, data = rm_total)

plot(allEffects(rm_total.mod.order), ylab = "Frequency of order", xlab = "Type of customer", main = NULL)
plot(allEffects(rm_total.mod.value), ylab = "Value of order", xlab = "Type of customer", main = NULL)

#descriptive differences of EATER_TYPE groups
ID_churn<- rm_total%>%
  filter(EATER_TYPE=="churn")%>%
  group_by(SUBSCRIBER_ID)

ID_retention<- rm_total%>%
  filter(EATER_TYPE=="retention")%>%
  group_by(SUBSCRIBER_ID)

percent_churn<-nrow(ID_churn)/(nrow(ID_churn)+nrow(ID_retention))

```

```{r echo=FALSE}
plot(allEffects(rm_total.mod.order),ylab = "Frequency of order", xlab = "Type of customer", main = NULL)
plot(allEffects(rm_total.mod.value),ylab = "Value of order", xlab = "Type of customer", main = NULL)
```

When further investigating the two groups we can see that the majority of customers belong to the Churn group (`r round(percent_churn,2)*100`% of all customers). Despite being smaller, the Retention group represents very valuable customers as they place more frequent and valuable orders than the Churn group (see figure below).

```{r include=FALSE}

#usuful plots for communicating that while the majoroty of the eaters did not stay in the subscription
#programme, RETENTION eaters are very valuable as they make a chunk of he order values and order numbers

#summay stats for EATER TYPES on ORDER value and number
sum_orders <- summarySE(rm_total, measurevar = "ORDERS", groupvars="EATER_TYPE")
sum_value <- summarySE(rm_total, measurevar = "ORDERS_VALUE", groupvars="EATER_TYPE")

#plot descriptives
g <- ggplot(rm_total, aes(EATER_TYPE,fill=EATER_TYPE))
g + geom_bar()

```

```{r echo=FALSE, include=FALSE}
ggplot(sum_orders, aes(x=EATER_TYPE, y=ORDERS, fill=EATER_TYPE)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=ORDERS-ci, ymax=ORDERS+ci),
              width=.2,               # Width of the error bars
              position=position_dodge(.9))+theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))+xlab("Type of customer")+ ylab("Average number of orders")
                           
ggplot(sum_value, aes(x=EATER_TYPE, y=ORDERS_VALUE, fill=EATER_TYPE)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=ORDERS_VALUE-ci, ymax=ORDERS_VALUE+ci),
              width=.2,               # Width of the error bars
              position=position_dodge(.9))+theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))+xlab("Type of customer")+ ylab("Average order value in £")
```

```{r echo=FALSE}

#summary stats for EATER TYPES on the number of ORDERS for different stages in the explored 10 month
sum_orders_stages <- summarySE(rm_total, measurevar = "ORDERS", groupvars=c("EATER_TYPE","SUBSCRIPTION_STATUS_IN_MONTH"))
#plot descriptives
ggplot(sum_orders_stages, aes(x=EATER_TYPE, y=ORDERS, fill=SUBSCRIPTION_STATUS_IN_MONTH)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=ORDERS-ci, ymax=ORDERS+ci),
              width=.2,               # Width of the error bars
              position=position_dodge(.9))+theme(panel.background = element_blank(),axis.line = element_line(colour = "black"))+xlab("Eater type")+ ylab("Average number of orders")+guides(fill = guide_legend(title = "Status"))

#summary stats for EATER TYPES on ORDERS_VALUE for different stages in the explored 10 month
sum_value_stages <- summarySE(rm_total, measurevar = "ORDERS_VALUE", groupvars=c("EATER_TYPE","SUBSCRIPTION_STATUS_IN_MONTH"))

ggplot(sum_value_stages, aes(x=EATER_TYPE, y=ORDERS_VALUE, fill=SUBSCRIPTION_STATUS_IN_MONTH)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=ORDERS_VALUE-ci, ymax=ORDERS_VALUE+ci),
              width=.2,               # Width of the error bars
              position=position_dodge(.9))+theme(panel.background = element_blank(),axis.line = element_line(colour = "black"))+xlab("Eater type")+ ylab("Average order value in £")+guides(fill = guide_legend(title = "Status"))

```

We can look at the differences in terms of generated revenue between the two groups (see table below). For the calculations below we used £2.5 as the delivery fee payerd by the customer, £7.99 as the subscription fee and 20% comission rate. Based on the tables below it is clear that the subscription programme generates increased revenue during the active and cancelled states, and the Retention group customers are outstandingly valuable for the business (economically speaking).

```{r include=FALSE}

revenue_table_churn<-data_frame(row.names = c("Delivery fee", "Subscription fee", "Comission"), pre_subscription = c(sum_orders_stages[1,4]*2.5,0,sum_value_stages[1,4]*0.2*sum_orders_stages[1,4]), cancelled = c(sum_orders_stages[2,4]*2.5,0,sum_value_stages[2,4]*0.2*sum_orders_stages[2,4]))

revenue_table_retention<-data_frame(row.names = c("Delivery fee", "Subscription fee", "Comission"), pre_subscription = c(sum_orders_stages[3,4]*2.5,0,sum_value_stages[3,4]*0.2*sum_orders_stages[3,4]), active = c(0,7.99,sum_value_stages[4,4]*0.2*sum_orders_stages[4,4]), cancelled = c(sum_orders_stages[5,4]*2.5,0,sum_value_stages[5,4]*0.2*sum_orders_stages[5,4]))
```
```{r echo=FALSE}

knitr::kable(revenue_table_retention, caption = "Summary of the average monthly generated revenue in the Retention group", col.names = c("revenue source","pre-subscription", "active", "cancelled"), digits=2)
```


```{r echo=FALSE}

knitr::kable(revenue_table_churn, caption = "Summary of the average monthly generated revenue in the Churn group", col.names = c("revenue source","pre-subscription", "cancelled"), digits=2)
```

```{r include=FALSE}
#Making some assumptions about the delivery associated cost for the business we can also calculate the profit generated by the two groups.
#Food delivery drivers on an average can do 3–4 deliveries in an hour.
#cost of rider £6-£7/ hour
#every succesful delivery earns the rider £1 
#estimated cost/order £6.5/£3.5 =  £1.86+£1 (for succesful order) 

#calculate costs of each EATER TYPE based on above estimations
cost_table_churn <- data_frame(row.names = c("Cost/order", "Riders_wage/order","Total cost/rider"), pre_subscription = c(sum_orders_stages[1,4]*1.86,sum_orders_stages[1,4]*1,sum_orders_stages[1,4]*2.86 ), cancelled = c(sum_orders_stages[2,4]*1.86,sum_orders_stages[2,4]*1,sum_orders_stages[1,4]*2.86 ))

cost_table_retention <- data_frame(row.names = c("Cost/order", "Riders_wage/order","Total cost/rider"), pre_subscription = c(sum_orders_stages[3,4]*1.86,sum_orders_stages[3,4]*1,sum_orders_stages[3,4]*2.86 ), active = c(sum_orders_stages[3,4]*1.86,sum_orders_stages[3,4]*1,sum_orders_stages[3,4]*2.86 ), cancelled = c(sum_orders_stages[5,4]*1.86,sum_orders_stages[5,4]*1,sum_orders_stages[5,4]*2.86))

#estimated summed profit for different EATER TYPES
colSums(revenue_table_churn[,-1])-cost_table_churn[3,2:3]
colSums(revenue_table_retention[,-1])-cost_table_retention[3,2:4]
```

As the Retention group proved to be valuable, it is worth gaining a deeper understanding on the behaviour of this group. For instance, it can be interesting to test whether having an active subscription changes their ordering behaviour in the *long-term*. To see whether increasing number of months have a growing facilitation effect as time passes I created a variable coding for the accumulated number of active monthly subscriptions. After this, linear regressions were run where the months passed was coded both as a continuous and as a categorical predictor. While the first one is good for quantifying general effect of time the other analysis can help to identify break-points within the 10-month period.

```{r include=FALSE}
#Let's get to know the RETAINMENT GROUP, any interesting features? Does having a subsciption change their 
#ordering behaviour in the long term?
#prepare dataset for TIME analysis, fisrt order dataset by DATE of the measures 
ordered_rm_ret_eaters_month<-rm_ret_eaters_month[order(rm_ret_eaters_month$SUBSCRIBER_ID,rm_ret_eaters_month$MONTH_START),]

ordered_rm_ret_eaters_month$MONTHS <- 1
end_point<- nrow(ordered_rm_ret_eaters_month)-1

#To be able to test TIME(i.e. MONTHS) effects we need to change DATE format to values and save it as a new
#variable

for (i in 1:end_point){
  if (ordered_rm_ret_eaters_month$MONTH_START[[i]]<ordered_rm_ret_eaters_month$MONTH_START[[i+1]])
  ordered_rm_ret_eaters_month$MONTHS[i+1] <- ordered_rm_ret_eaters_month$MONTHS[i]+1
}

#we can have MONTHS both as a categorical (useful to see shifts between months) and as continious variable
#(useful for talking about general time effects)
#continous variable needs to be centered first
ordered_rm_ret_eaters_month$Cont_MONTHS<-scale(ordered_rm_ret_eaters_month$MONTHS,center = TRUE, scale = TRUE)
#categorical variable declared as a factor
ordered_rm_ret_eaters_month$MONTHS<-factor(ordered_rm_ret_eaters_month$MONTHS)
#run model MONTHS as categorical for number of ORDERS
time_ret.mod.order<-lm(ORDERS ~  MONTHS, data = ordered_rm_ret_eaters_month)
#gather stats
summary(time_ret.mod.order)
#plot estimates
plot(allEffects(time_ret.mod.order))
#run model MONTHS as categorical for number of ORDERS
time_ret.mod.value<-lm(ORDERS_VALUE ~  MONTHS, data = ordered_rm_ret_eaters_month)
#gather stats
summary(time_ret.mod.value)
#plot estimates
plot(allEffects(time_ret.mod.value)) 
# interesting change after the 3rd month, 10th month is probably noise
# maybe dataset was generated at the beginning of the 10th month

#run model MONTHS as continous for number of ORDERS
ctime_ret.mod.order<-lm(ORDERS ~  Cont_MONTHS, data = ordered_rm_ret_eaters_month)
#gather stats
summary(ctime_ret.mod.order)
#plot estimates
plot(allEffects(ctime_ret.mod.order))

ctime_ret.mod.value<-lm(ORDERS_VALUE ~  Cont_MONTHS, data = ordered_rm_ret_eaters_month)
#gather stats
summary(ctime_ret.mod.value)
#plot estimates
plot(allEffects(ctime_ret.mod.value))
```

The analysis revealed a significant positive linear relationship between the increasing number of active subscription months and increasing number of order frequency and value. When the active subscription months were coded as categories one can see the significant increases appear from the fourth active month. While in the tenth month there seems to be a significant drop it is likely to be arbitrary, probably because the data collection was finished at beginning of the tenth (final) month (see figures below). 

```{r echo=FALSE}
plot(allEffects(ctime_ret.mod.order),ylab = "Frequency of order", xlab = "Number of active months",main = "Retention group")
plot(allEffects(ctime_ret.mod.value),ylab = "Value of order", xlab = "Number of active months",main = "Retention group")
```

```{r echo=FALSE}
plot(allEffects(time_ret.mod.order),ylab = "Frequency of order", xlab = "Number of active months",main = "Retention group")
plot(allEffects(time_ret.mod.value),ylab = "Value of order", xlab = "Number of active months",main = "Retention group") 

```

## Limitations and future steps

While the statistical tests provide confidence in the accuracy of these results it must be noted that the some of the assumptions these tests hold are violated by the distribution of the analysed sample. Accuracy could be further improved by acquiring exact data on monthly costs and revenue associated for every customer. Furthermore, collaboration with the product management and user research team would help with the interpretation and validation of the results regarding different types of subscribers. Finally, data on the customers' activity from the business's competitors would deepen the insights on the impact of the subscription programme.       



